---
title: "Stat 1 R Code - Week 10"
author: "SMU"
date: "4/21/2022"
output: html_document
---

# Cleaning the global environment
```{r}
rm(list = ls())
```

```{r include = FALSE}
library(tidyverse)
library(ggplot2)
library(scales)
library(pwr)
library(agricolae)
library(huxtable)
library(lawstat)
library(lsmeans)
library(nCDunnett)
library(dplyr)
library(investr)
library(multcomp)
library(pairwiseCI)
library(DescTools)
```



# Example

```{r}
Data = data.frame( score = c(3,5,7,10,12,14,20,22,24),  level = c(1,1,1,2,2,2,3,3,3))
fit = lm(score ~level, data = Data)
anova(fit)
summary(fit)
```

# Study Hour Example
```{r}
Data1 = data.frame( examscore = c( 34, 56, 145, 270,255, 268, 367, 479, 445, 489, 695, 778, 794), studyhours = c(1,1,2,2,2,3,4,4,4,6,7,7,8) )
fit = lm(examscore ~studyhours, data = Data1)
anova(fit)
summary(fit)
```

# Confidence Interval
```{r}
predict(fit, interval = "confidence", level = 0.95)
```
# Confidence interval for the mmean of Y when the X is not in the original dataset

```{r}
newpoint = data.frame(examscore = NA, studyhours = 5)
predict(fit, newdata = newpoint, interval = "confidence")
```

# Prediction Interval
```{r}
predict(fit, interval = "prediction", level = 0.95)
```


# Prediction interval for the mmean of Y when the X is not in the original dataset

```{r}
newpoint = data.frame(examscore = NA, studyhours = 5)
predict(fit, newdata = newpoint, interval = "prediction")
```

# Calibrate interval for the mean grade
```{r}
calibrate(fit, y0 = 75, interval = "Wald", mean.response = TRUE, limit = FALSE)
```
# Calibrate interval for an individual grade
```{r}
calibrate(fit, y0 = 75, interval = "Wald", mean.response = FALSE, limit = FALSE)
```


# Visualizing Data before Regression
```{r}
#confidence and prediction intervals
movies = read.csv("D:/SMU/Ph.D./Data Science/Statistical Foundation for Data Science/Assignment/Unit 10/Male Display Data Set.csv")
movies.lm=lm(Tcell~Mass, data = movies)
newx=movies$Mass
newx=sort(newx)

prd_c=predict(movies.lm, newdata= data.frame(Mass = newx), interval=c("confidence"), type = c("response"), level=0.99) 
prd_p=predict(movies.lm, newdata= data.frame(Mass = newx), interval=c("prediction"), type = c("response"), level=0.99) 


#Plot with confidence and prediction intervals
plot(movies[,1],movies[,2],xlim = c(3,9), ylim = c(-0.1,0.8),xlab = "Mass",ylab = "Tcell", main = "Tcell versus Mass")
abline(movies.lm, col = "red")
lines(newx,prd_c[,2],col = "blue",lty = 2, lwd = 2)
lines(newx,prd_c[,3],col = "blue", lty = 2, lwd = 2)
lines(newx,prd_p[,2],col = "green", lty = 2, lwd = 2)
lines(newx,prd_p[,3],col = "green", lty = 2, lwd = 2)

```

# HW 10

```{r}
movies = read.csv("D:/SMU/Ph.D/Data Science/Assignment/Unit 10/Male Display Data Set.csv")
movies.lm=lm(Tcell~Mass, data = movies)
newx=movies$Mass
newx=sort(newx)
prd_c=predict(movies.lm, newdata= data.frame(Mass = newx), interval=c("confidence"), type = c("response"), level=0.99) 
prd_p=predict(movies.lm, newdata= data.frame(Mass = newx), interval=c("prediction"), type = c("response"), level=0.99) 
#Plot with confidence and prediction intervals
plot(movies[,1],movies[,2],xlim = c(3,9), ylim = c(-0.1,0.8),xlab = "Mass",ylab = "Tcell", main = "Tcell versus Mass")
abline(movies.lm, col = "red")
lines(newx,prd_c[,2],col = "blue",lty = 2, lwd = 2)
lines(newx,prd_c[,3],col = "blue", lty = 2, lwd = 2)
lines(newx,prd_p[,2],col = "green", lty = 2, lwd = 2)
lines(newx,prd_p[,3],col = "green", lty = 2, lwd = 2)
```

```{r}
movies.lm=lm(Tcell~Mass, data = movies)
summary(movies.lm)
```
# Confidence interval for the mmean of Y when the X is not in the original dataset

```{r}
newpoint = data.frame(Tcell = NA, Mass = 4.5)
predict(movies.lm, newdata = newpoint, interval = "confidence")
```

# Prediction interval for the mmean of Y when the X is not in the original dataset

```{r}
newpoint = data.frame(Tcell = NA, Mass = 4.5)
predict(movies.lm, newdata = newpoint, interval = "prediction")
```


# Visualizing Data before Regression
```{r}
#confidence and prediction intervals
movies = read.csv("D:/SMU/Data Science/Assignment/Unit 10/Male Display Data Set.csv")
movies.lm=lm(Tcell~Mass, data = movies)
newx=movies$Mass
newx=sort(newx)
prd_c=predict(movies.lm, newdata= data.frame(Mass = newx), interval=c("confidence"), type = c("response"), level=0.99) 
prd_p=predict(movies.lm, newdata= data.frame(Mass = newx), interval=c("prediction"), type = c("response"), level=0.99) 
#Plot with confidence and prediction intervals
plot(movies[,1],movies[,2],xlim = c(2,11), ylim = c(-0.1,0.8),xlab = "Mass",ylab = "Tcell", main = "Tcell versus Mass")
abline(movies.lm, col = "red")
lines(newx,prd_c[,2],col = "blue",lty = 2, lwd = 2)
lines(newx,prd_c[,3],col = "blue", lty = 2, lwd = 2)
lines(newx,prd_p[,2],col = "green", lty = 2, lwd = 2)
lines(newx,prd_p[,3],col = "green", lty = 2, lwd = 2)
abline(h = 0.3, col = "red", lwd = 4)
#abline(v = 115, col = "green", lwd = 3)

```

# Calibrate interval for the mean t-cell
```{r}
calibrate(movies.lm, y0 = 0.3, level = 0.99, interval = "Wald", mean.response = TRUE, limit = FALSE)
```

# Calibrate interval for an individual t-cell
```{r}
calibrate(movies.lm, y0 = 0.3, level = 0.99, interval = "Wald", mean.response = FALSE, limit = FALSE)
```

# Residual
```{r}
plot(movies.lm$residuals, main="Scatterplot",
   xlab="Predicted Values", ylab="Residuals", pch=19)
```

# Residual Histogram with Normal Distribution
```{r}
g = movies.lm$residuals
m<-mean(g)
std<-sqrt(var(g))
hist(g, density=20, breaks=20, prob=TRUE, col="red",
     xlab="Residuals", ylim=c(0, 8), 
     main="Residual Histogram with Normal Distribution")
curve(dnorm(x, mean=m, sd=std), 
      col="darkblue", lwd=2, add=TRUE, yaxt="n")
```
# R-Squared
```{r}
summary(movies.lm)
```

